---
title: "Gastric Cancer Dataset Practise - GSE62254"
author: "Natasha Windmeyer"
date: "19/03/2021"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This Practise document is using the Affymetrix microarray dataset (GSE62254) - re-normalised using RMA without background correction

```{r}
install.packages('here')
```

```{r}
library(here)
library(dplyr)
library(ggplot2)
```
The 'here()' starts at /Users/black/GitHub/Gastric_Cancer_Datasets

```{r}
options(width=80)
```
This sets the output wider to 80, making things look a bit nicer

```{r}
load(here('Data/gse62254_gastric_cancer.RData'))
```
This loads the GSE62254 Gastric Cancer Dataset

```{r}
ls()
```
This confirms what was loaded

```{r}
class(gse62254_clinDat)
class(gse62254_expDat)
```
This tells you what type of objects the datasets are

```{r}
dim(gse62254_clinDat)
dim(gse62254_expDat)
```
This gives you the dimensions of the datasets

The Variable names for the Clinical Dataset
```{r}
names(gse62254_clinDat)
```
dfsEvent: Disease Free Survival Event (0,1)
dfsTime: Time of Disease Free Survival Event (months)
lc: Lauren Classification
molSub: Molecular Subtype (Cristescu et al)
stage: Tumour stage
region: Anatomic region of tumour
gender: Patient gender
ageCat: Patient age category

```{r}
head(gse62254_clinDat)
```
Gives you the first 6 lines of the clinical dataset

The sample names are not included in the clinical data, so we need to add this as a new column
```{r}
gse62254_clinDat = gse62254_clinDat %>% mutate(, PatID = colnames(gse62254_expDat))
head(gse62254_clinDat)
```

